Performance Models for Data Transfers: A Case Study with Molecular Chemistry Kernels

### 2019-04-21 ICPP-19/communications.tex
attempt to fit the document in 10 pages

	The outline of the paper is the following. Section~\ref{sec:relatedWork} describes past work on the computations with limited memory and similar problems in the literature. In section~\ref{sec:theoreticalProof}, we present an algorithm to obtain the order of data transfers when there is not any memory capacity restriction. Then, we also prove that in general data transfer problem is NP-complete. In Section~\ref{sec:heuristics}, we propose several heuristics and discuss the conditions that might be favorable to them. We mainly consider three categories of heuristics: static heuristics, dynamic heuristics and static heuristics with dynamic corrections. Sections~\ref{sec:expSetting} describes our experimental setup and we evaluate our proposed strategies on two molecular chemistry kernels in Section~\ref{sec:expResults}. Our results show that static heuristics with dynamic corrections achieve good performance in most cases. We finally propose conclusions and perspectives in Section~\ref{sec:conclusion}.

@@

	\noindent\textbf{Problem $DT$} : A set of tasks $ST=\{T_1,
	\cdots, T_n\}$ is scheduled on a processing unit $P$ with
	memory unit $M$ of capacity $C$. Input data for tasks of $ST$
	reside on another memory unit $M'$. $COMM_i$ is the communication time to
	transfer input data from $M'$ to $M$ for task $i$ and $COMP_i$
	is the computation time of task $i$ on $P$. We assume that these
	tasks do not produce any output data. There can be only one
	communication at a time, and $P$ can only process one task at
	a time. A task uses an amount of memory in  $M$ from the
	start of its communication to the end of its computation.

@@

	\begin{algorithm}
		\caption{\label{alg:OrderOfExecutionInfinteMemory}Johnson's~\cite{johnson} algorithm (infinite memory case).}
		\begin{algorithmic}[1]
			\STATE Divide ready tasks in two sets $S_1$ and $S_2$. If computation time of a task $T$ is not less than its communication time, then $T$ is in $S_1$ otherwise in $S_2$.
			\STATE Sort $S_1$ in queue $Q$ by non-decreasing communication times
			\STATE Sort $S_2$ in queue $Q'$ by non-increasing computation times
			\STATE Append $Q'$ to $Q$
			\STATE $\tau_{\text{COMM}} \gets 0$ \hfill\COMMENT{Available time of communication resource}
			\STATE $\tau_{\text{COMP}} \gets 0$\hfill \COMMENT{Available time of computation resource}
			\WHILE{$Q \neq \emptyset$}
			\STATE Remove a task $T$ from beginning of $Q$ for processing
			\STATE $\scomm(T) \gets \tau_{\text{COMM}}$
			\STATE $\scomp(T) \gets max(\scomm(T) + COMM_T, \tau_{\text{COMP}})$
			\STATE $\tau_{\text{COMM}} \gets \scomm(T) + COMM_T$
			\STATE $\tau_{\text{COMP}} \gets \scomp(T) + COMP_T$
			\ENDWHILE
		\end{algorithmic}
	\end{algorithm}
	\vspace*{-0.35cm}
	We also prove optimality of Algorithm~\ref{alg:OrderOfExecutionInfinteMemory} differently in an extended version~\cite{extendedVersion}.
	\subsection{Finite Memory}

@@

	\begin{table}[!ht]
		\begin{center}
			\begin{tabular}{|c|c|c|c|}
				\hline
				%%				Task & Comm Volume & Comm Time & Comp Time\\ \hline 
				\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
				&=Comm Volume && \\ \hline
				A & 0 & 0 & 5\\ \hline
				B & 4 & 4 & 3\\ \hline
				C & 1 & 1 & 6\\ \hline
				D & 3 & 3 & 7\\ \hline
				E & 6 & 6 & 0.5\\ \hline
				F & 7 & 7 & 0.5\\ \hline
			\end{tabular}
		\end{center}
		\caption{Example instance where ordering on both resources
			has to be different\vspace*{-0.9cm}.}
		\label{table:different.order}
	\end{table}

@@

	We classify our heuristics into mainly three categories. In the first category, the order of all computations and communications is computed in advance and the same order is followed on both resources. In the second category, the next task to schedule is dynamically chosen based on different criteria. The final category is based on combining strategies from the first two categories: a static ordering is precomputed and corrected dynamically to avoid idle time caused by memory limitations. In all of our strategies (except linear programming based strategy), communication and computations take place in the same order.

@@

\begin{figure}[!tb]
	\newcommand{\taskA}[2][0]{\task[#1]{$A$}{#2}{3}{2}{cyan}}
	\newcommand{\taskB}[2][0]{\task[#1]{$B$}{#2}{1}{3}{blue!40!white}}
	\newcommand{\taskC}[2][0]{\task[#1]{$C$}{#2}{4}{4}{blue!70!white}}
	\newcommand{\taskD}[2][0]{\task[#1]{$D$}{#2}{2}{1}{blue}}
\vspace*{-0.25cm}	
	\centering
	\subfloat[][Infinite Memory Capacity]{
		\begin{tikzpicture}[scale=0.6]
		\schedule{12.5}{OMIM}{1,4,5,8,9,10,11,12}
		\taskB{0}
		\taskC{1}
		\taskA[1]{5}
		\taskD[1]{8}
		\end{tikzpicture}
	}

	\subfloat[][Memory Capacity: 6]{
		\begin{tikzpicture}[yscale=0.6, xscale=0.45]
		\begin{scope}
		\schedule{15.5}{OOSIM}{1,4,5,9,12,14,15}
		\taskB{0}
		\taskC{1}
		\taskA{9}
		\taskD{12}
		\end{scope}
		\begin{scope}[yshift=-2.75cm]
		\schedule{16.5}{IOCMS}{1,3,4,5,6,8,12,16}
		\taskB{0}
		\taskD[1]{1}
		\taskA{3}
		\taskC{8}
		\end{scope}
		\begin{scope}[yshift=-5.5cm]
		\schedule{14.5}{DOCPS}{4,5,8,11,13,14}
		\taskC{0}
		\taskB[3]{4}
		\taskA{8}
		\taskD{11}
		\end{scope}
		\begin{scope}[yshift=-8.25cm]
		\schedule{16.5}{IOCCS}{2,3,6,8,12,16}
		\taskD{0}
		\taskB{2}
		\taskA{3}
		\taskC{8}
		\end{scope}
		\begin{scope}[yshift=-11cm]
		\schedule{17.5}{DOCCS}{4,8,11,12,13,14,16,17}
		\taskC{0}
		\taskA{8}
		\taskB[1]{11}
		\taskD[2]{12}
		\end{scope}
		\end{tikzpicture}
	}

@@

	\begin{table}[!tb]
	\begin{center}
%%\vspace*{-0.10cm}		
		\begin{tabular}{|c|c|c|c|}
			\hline
			%%				\multirow{2}{*}{Task} & Comm Time & \multirow{2}{*}{Comp Time}\\ 
			%%				&=Comm Volume& \\ \hline
			%%				Task & Comm Volume & Comm Time & Comp Time\\ \hline 
			\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
			&=Comm Volume && \\ \hline
			A & 3 & 3 &  2\\ \hline
			B & 1 & 1 & 3\\ \hline
			C & 4 & 4 & 4\\ \hline
			D & 2 & 2 & 1\\ \hline
		\end{tabular}
		\caption{\label{tab:staticOrderExample} A task set for static order schedules.\vspace*{-0.75cm}}
	\end{center}
\end{table}

@@

		\centering
		\begin{tikzpicture}[yscale=0.6,xscale=0.3]
		\begin{scope}
		\schedule{24}{LCMR}{1,6,7,8,11,13,17,23}
		\taskB{0}
		\taskD[1]{1}
		\taskA{8}
		\taskC{13}
		\end{scope}
		\begin{scope}[yshift=-2.75cm]
		\schedule{26}{SCMR}{1,4,7,9,13,19,24,25}
		\taskB{0}
		\taskA[3]{1}
		\taskC{9}
		\taskD{19}
		\end{scope}
		\begin{scope}[yshift=-5.5cm]
		\schedule{25}{MAMR}{1,5,7,13,16,18,23,24}
		\taskB{0}
		\taskC[2]{1}
		\taskA{13}
		\taskD{18}
		\end{scope}
		\end{tikzpicture}
		\caption{ \label{fig:dynamicSelectionExample} Different dynamic heuristic schedules for a task set of Table~\ref{tab:dynamicSelectionExample} with a memory capacity of 6.\vspace*{-0.25cm}}
		%%		 Optimal schedule length is 23.
	\end{figure}

	\begin{table}[!tb]
	\begin{center}
		\begin{tabular}{|c|c|c|c|}
			\hline
			%%				Task & Comm Volume & Comm Time & Comp Time\\ \hline 
			\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
			&=Comm Volume && \\ \hline
			A & 3 & 3 & 2\\ \hline
			B & 1 & 1 &  6\\ \hline
			C & 4 & 4 & 6\\ \hline
			D & 5 & 5 & 1\\ \hline
		\end{tabular}
		\caption{\label{tab:dynamicSelectionExample} A task set for dynamic schedules.\vspace*{-0.5cm}}
	\end{center}
	\end{table}

@@

	\subsection{Static Order with Dynamic Corrections}

	In this class of strategy, we precompute the order of tasks based on some criterion and then follow this ordering as much as possible. But when the communication resource is idle because the memory requirement of the next task is too high, then we select a task  with a dynamic strategy. After a task is dynamically selected, we update the remaining order without this task. This class of strategy takes advantage of static information in the form of precomputed order and dynamic corrections to minimize the idle time due to memory constraint.

@@

	%%\begin{figure}[htb]
	%%\includegraphics[scale=0.05]{Figs/staticOrderDynamicCorrectionsSchedules}
	%%\caption{Task set and schedules where orders are precomputed based on different heuristics.}
	%%\label{fig:staticOrderDynamicCorrectionsExample}
	%%\end{figure}
\begin{figure}[!htb]
	\newcommand{\taskA}[2][0]{\task[#1]{$A$}{#2}{4}{1}{cyan}}
	\newcommand{\taskB}[2][0]{\task[#1]{$B$}{#2}{2}{6}{cyan!50!black}}
	\newcommand{\taskC}[2][0]{\task[#1]{$C$}{#2}{8}{8}{blue!40!white}}
	\newcommand{\taskD}[2][0]{\task[#1]{$D$}{#2}{5}{4}{blue!70!white}}
	\newcommand{\taskE}[2][0]{\task[#1]{$E$}{#2}{3}{2}{blue}}
	\begin{tikzpicture}[yscale=0.6,xscale=0.2]
	\begin{scope}
	\schedule{35}{OOLCMR}{2,8,12,15,17,25,33}
	\taskB{0}
	\taskD[1]{2}
	\taskA{8}
	\taskE{12}
	\taskC{17}
	\end{scope}
	\begin{scope}[yshift=-2.75cm]
	%%		\schedule{36}{OOSCMR}{2,5,8,11,14,18,26,34}
	%%		\taskB{0}
	%%		\taskE[3]{2}
	%%		\taskA[1]{5}
	%%		\taskD{9}
	%%		\taskC{18}
	%%		\end{scope}
	\schedule{36}{OOSCMR}{2,5,8,11,15,19,27,35}
	\taskB{0}
	\taskE[3]{2}
	\taskA[1]{5}
	\taskD{10}
	\taskC{19}
	\end{scope}
	\begin{scope}[yshift=-5.5cm]
	\schedule{36}{OOMAMR}{2,8,11,14,17,25,33}
	\taskB{0}
	\taskD[1]{2}
	\taskE[1]{8}
	\taskA{12}
	\taskC{17}
	\end{scope}
	%%		\begin{scope}[yshift=+2.75cm]
	%%		\schedule{40}{OMIM}{2,8,16,24,29,33,38}
	%%		\taskB{0}
	%%		\taskC{8}
	%%		\taskD{24}
	%%		\taskA{29}
	%%		\taskE{33}
	%%		\end{scope}

	\end{tikzpicture}
	\caption{ \label{fig:staticOrderDynamicCorrectionsExample} Different static order dynamic corrections heuristic schedules for a task set of Table~\ref{tab:staticOrderDynamicCorrectionsExample} with a memory capacity of 9. The $OMIM$ order is $BCDAE$.\vspace*{-0.5cm}}
\end{figure}

\begin{table}[!b]
	\begin{center}
		\begin{tabular}{|c|c|c|c|}
			\hline
			%%				Task & Comm Volume & Comm Time & Comp Time\\ \hline 
			\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
			&=Comm Volume && \\ \hline
			A & 4 & 4 &  1\\ \hline
			B & 2 & 2 & 6\\ \hline
			C & 8 & 8 & 8\\ \hline
			D & 5 & 5 & 4\\ \hline
			E & 3 & 3 & 2\\ \hline
		\end{tabular}
		\caption{\label{tab:staticOrderDynamicCorrectionsExample} A task set for static order dynamic corrections schedules.}
	\end{center}
\end{table}

	As previously, we propose on Table~\ref{tab:staticOrderDynamicCorrectionsExample} an example instance for this class of strategies, and provide on Figure~\ref{fig:staticOrderDynamicCorrectionsExample} the corresponding schedules.

@@

		\begin{table}[!h]
			\begin{tabular}{|c|p{6.5cm}|}
				\hline
				\textbf{Heuristic} & \textbf{\hspace{2cm}Favorable Situation} \\ \hline
				$OOSIM$ & Memory capacity is not a restriction (\textcolor{green}{Optimal}) \\ \hline
				$IOCMS$ & Memory capacity is not a restriction and tasks are compute intensive (\textcolor{green}{Optimal}) \\ \hline
				$DOCPS$ & Memory capacity is not a restriction and tasks are communication intensive (\textcolor{green}{Optimal}) \\ \hline
				$IOCCS$ & Moderate memory capacity and most tasks are highly compute intensive \\ \hline
				$DOCCS$ & Moderate memory capacity and most tasks are highly communication intensive \\ \hline
				$LCMR$ & Limited memory capacity and significant percentage of tasks with large communication times are compute intensive\\ \hline
				$SCMR$ & Limited memory capacity and significant percentage of tasks with small communication times are compute intensive\\ \hline
				$MAMR$ & Limited memory capacity and significant percentage of all types of tasks\\ \hline
				$OOLCMR$ & Moderate memory capacity and significant percentage of slightly communication intensive tasks have large communication times\\ \hline
				$OOSCMR$ & Moderate memory capacity and significant percentage of slightly communication intensive tasks have small communication times\\ \hline
				$OOMAMR$ & Moderate memory capacity and significant percentage of all types of tasks \\ \hline
			\end{tabular}\caption{~\label{tab:heuristicsAndFavorableScenarios}Heuristics and their favorable scenarios.}
		\end{table}
			Based on the definition of proposed strategies and the optimality of Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}, we discuss the scenarios which might be more favorable for each heuristic in Table~\ref{tab:heuristicsAndFavorableScenarios}. This allows programmers to use appropriate strategies to maximize communication-computation overlap for their applications. In this table, ``moderate memory capacity'' refers to the case where memory is constrained, but close to the maximal memory requirement of the $OMIM$ schedule.

@@

		%%		\todo[inline]{More about tiling}
		Both applications mainly perform two types of computations, tensor transpose and tensor contraction. HF expects to specify a tile size and we set it to 100, so that each core can be used efficiently. CCSD automatically determines tile sizes at different program points based on the input molecules. Hence, HF operates on almost homogeneous tiles while CCSD uses more heterogeneous tiles.
\begin{figure*}[!t]
	\includegraphics[scale=0.5]{./results/ratio_to_optimal_selected_hf.pdf}
	\vspace*{-0.85cm}\caption{Comparison of different heuristics for HF with $m_c=176KB$.			\vspace*{-0.35cm}}
	\label{fig:ratio_to_optimal_hf}
\end{figure*}			
		\subsection{Workload Characteristics}

		\begin{figure}[htb]
\begin{center}
			\includegraphics[scale=0.48]{./results/application_properties.pdf}
			\vspace*{-0.9cm}\caption{HF and CCSD tasks characteristics.\vspace*{-0.5cm}}
			\label{fig:ApplicationProperties}
\end{center}
		\end{figure}

		To get more insights into the considered workloads, we provide  information about the instances we consider in Figure~\ref{fig:ApplicationProperties}. For each instance, we compute the sum of communication times (\textit{sum comm}) and the sum of computation times  (\textit{sum comp}), and normalize it relatively to the $OMIM$ value. Figure~\ref{fig:ApplicationProperties} also shows the maximum of both values, which is a lower bound on the possible makespan of a schedule for this instance, and their sum, which is an upper bound: this represents the makespan of the sequential schedule, obtained with zero overlap between computation and communication. We can see that HF is a communication intensive application and at most 20\% overlap can be expected in the best scenario. On the other hand, in the CCSD workload, communications and computations are almost evenly distributed and a more significant overlap is possible.

@@

		The CCSD application operates on tasks of different sizes, hence different heuristics exhibit distinct behaviors even at minimum memory capacity $m_c$. Heterogeneity favors dynamic strategies, as can be seen by the fact that both dynamic and static order with dynamic corrections based strategies perform better than static based strategies. Similar to HF, static order with dynamic corrections based strategies outperform others as memory capacity becomes moderate.

		\begin{figure}[!t]
			\includegraphics[scale=0.45]{./results/ratio_to_optimal_ccsd-best.pdf}
			\vspace*{-0.45cm}\caption{Best variants of all categories for CCSD.\vspace*{-0.45cm}}
			\label{fig:ratio_to_optimal_best_ccsd}
		\end{figure}

>>>>

	The outline of the paper is the following. Section~\ref{sec:relatedWork} describes past work on the computations with limited memory and similar problems in the literature. In section~\ref{sec:theoreticalProof}, we present an algorithm to obtain the order of data transfers when there is not any memory capacity restriction. Then, we also prove that in general data transfer problem is NP-complete. In Section~\ref{sec:heuristics}, we propose several heuristics and discuss the conditions that might be favorable to them. We mainly consider three categories of heuristics: static order based heuristics, dynamic selection based heuristics, and static order with dynamic correction based heuristics. Sections~\ref{sec:expSetting} describes our experimental setup and we evaluate our proposed strategies on two molecular chemistry kernels in Section~\ref{sec:expResults}. Our results show that static order with dynamic correction based heuristics achieve good performance in most cases. We finally propose conclusions and perspectives in Section~\ref{sec:conclusion}.

@@

	\medskip
	\noindent\textbf{Problem $DT$} : A set of tasks $ST=\{T_1,
	\cdots, T_n\}$ is scheduled on a processing unit $P$ with
	memory unit $M$ of capacity $C$. Input data for tasks of $ST$
	reside on another memory unit $M'$. $COMM_i$ is the communication time to
	transfer input data from $M'$ to $M$ for task $i$ and $COMP_i$
	is the computation time of task $i$ on $P$. We assume that these
	tasks do not produce any output data. There can be only one
	communication at a time, and $P$ can only process one task at
	a time. A task uses an amount of memory in  $M$ from the
	start of its communication to the end of its computation.

@@

	\vspace*{-0.25cm}
	\begin{algorithm}
		\caption{\label{alg:OrderOfExecutionInfinteMemory}Johnson's~\cite{johnson} algorithm (infinite memory case).}
		\begin{algorithmic}[1]
			\STATE Divide ready tasks in two sets $S_1$ and $S_2$. If computation time of a task $T$ is not less than its communication time, then $T$ is in $S_1$ otherwise in $S_2$.
			\STATE Sort $S_1$ in queue $Q$ by non-decreasing communication times
			\STATE Sort $S_2$ in queue $Q'$ by non-increasing computation times
			\STATE Append $Q'$ to $Q$
			\STATE $\tau_{\text{COMM}} \gets 0$ \hfill\COMMENT{Available time of communication resource}
			\STATE $\tau_{\text{COMP}} \gets 0$\hfill \COMMENT{Available time of computation resource}
			\WHILE{$Q \neq \emptyset$}
			\STATE Remove a task $T$ from beginning of $Q$ for processing
			\STATE $\scomm(T) \gets \tau_{\text{COMM}}$
			\STATE $\scomp(T) \gets max(\scomm(T) + COMM_T, \tau_{\text{COMP}})$
			\STATE $\tau_{\text{COMM}} \gets \scomm(T) + COMM_T$
			\STATE $\tau_{\text{COMP}} \gets \scomp(T) + COMP_T$
			\ENDWHILE
		\end{algorithmic}
	\end{algorithm}
	\vspace*{-0.35cm}
	We also prove optimality of Algorithm~\ref{alg:OrderOfExecutionInfinteMemory} differently in an extended version~\cite{extendedVersion}.
	\subsection{Finite Memory}

@@

	\begin{table}[!ht]
		\begin{center}
			\begin{tabular}{|c|c|c|c|}
				\hline
				Task & Memory Req & Comm Time & Comp Time\\ \hline 
%%				\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
%%				&=Comm Volume && \\ \hline
				A & 0 & 0 & 5\\ \hline
				B & 4 & 4 & 3\\ \hline
				C & 1 & 1 & 6\\ \hline
				D & 3 & 3 & 7\\ \hline
				E & 6 & 6 & 0.5\\ \hline
				F & 7 & 7 & 0.5\\ \hline
			\end{tabular}
		\end{center}
		\caption{Example instance where ordering on both resources
			has to be different\vspace*{-0.9cm}.}
		\label{table:different.order}
	\end{table}

@@

	We classify our heuristics into mainly three categories. In the first category, the order of all computations and communications is computed in advance and the same order is followed on both resources. In the second category, the next task to schedule is dynamically chosen based on different criteria. The final category is based on combining strategies from the first two categories: a static order is precomputed and corrected dynamically to avoid idle time caused by memory limitations. In all of our strategies (except linear programming based strategy), communication and computations take place in the same order.

@@

\begin{figure}[!tb]
	\newcommand{\taskA}[2][0]{\task[#1]{$A$}{#2}{3}{2}{cyan}}
	\newcommand{\taskB}[2][0]{\task[#1]{$B$}{#2}{1}{3}{blue!40!white}}
	\newcommand{\taskC}[2][0]{\task[#1]{$C$}{#2}{4}{4}{blue!70!white}}
	\newcommand{\taskD}[2][0]{\task[#1]{$D$}{#2}{2}{1}{blue}}
\vspace*{-0.30cm}	
	\centering
	\subfloat[][Infinite Memory Capacity]{
		\begin{tikzpicture}[yscale=0.5, xscale=0.6]
		\schedule{12.5}{OMIM}{1,4,5,8,9,10,11,12}
		\taskB{0}
		\taskC{1}
		\taskA[1]{5}
		\taskD[1]{8}
		\end{tikzpicture}
	}

	\subfloat[][Memory Capacity: 6]{
		\begin{tikzpicture}[yscale=0.5, xscale=0.45]
		\begin{scope}
		\schedule{15.5}{OOSIM}{1,4,5,9,12,14,15}
		\taskB{0}
		\taskC{1}
		\taskA{9}
		\taskD{12}
		\end{scope}
		\begin{scope}[yshift=-2.75cm]
		\schedule{16.5}{IOCMS}{1,3,4,5,6,8,12,16}
		\taskB{0}
		\taskD[1]{1}
		\taskA{3}
		\taskC{8}
		\end{scope}
		\begin{scope}[yshift=-5.5cm]
		\schedule{14.5}{DOCPS}{4,5,8,11,13,14}
		\taskC{0}
		\taskB[3]{4}
		\taskA{8}
		\taskD{11}
		\end{scope}
		\begin{scope}[yshift=-8.25cm]
		\schedule{16.5}{IOCCS}{2,3,6,8,12,16}
		\taskD{0}
		\taskB{2}
		\taskA{3}
		\taskC{8}
		\end{scope}
		\begin{scope}[yshift=-11cm]
		\schedule{17.5}{DOCCS}{4,8,11,12,13,14,16,17}
		\taskC{0}
		\taskA{8}
		\taskB[1]{11}
		\taskD[2]{12}
		\end{scope}
		\end{tikzpicture}
	}

@@

	\begin{table}[!tb]
	\begin{center}
%%\vspace*{-0.10cm}		
		\begin{tabular}{|c|c|c|c|}
			\hline
			%%				\multirow{2}{*}{Task} & Comm Time & \multirow{2}{*}{Comp Time}\\ 
			%%				&=Comm Volume& \\ \hline
	    	Task & Memory Req & Comm Time & Comp Time\\ \hline 
%%			\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
%%			&=Comm Volume && \\ \hline
			A & 3 & 3 &  2\\ \hline
			B & 1 & 1 & 3\\ \hline
			C & 4 & 4 & 4\\ \hline
			D & 2 & 2 & 1\\ \hline
		\end{tabular}
		\caption{\label{tab:staticOrderExample} A task set for static order schedules.\vspace*{-0.75cm}}
	\end{center}
\end{table}

@@

		\centering
		\begin{tikzpicture}[yscale=0.5,xscale=0.3]
		\begin{scope}
		\schedule{24}{LCMR}{1,6,7,8,11,13,17,23}
		\taskB{0}
		\taskD[1]{1}
		\taskA{8}
		\taskC{13}
		\end{scope}
		\begin{scope}[yshift=-2.75cm]
		\schedule{26}{SCMR}{1,4,7,9,13,19,24,25}
		\taskB{0}
		\taskA[3]{1}
		\taskC{9}
		\taskD{19}
		\end{scope}
		\begin{scope}[yshift=-5.5cm]
		\schedule{25}{MAMR}{1,5,7,13,16,18,23,24}
		\taskB{0}
		\taskC[2]{1}
		\taskA{13}
		\taskD{18}
		\end{scope}
		\end{tikzpicture}
		\caption{ \label{fig:dynamicSelectionExample} Different dynamic heuristic schedules for a task set of Table~\ref{tab:dynamicSelectionExample} with a memory capacity of 6.\vspace*{-0.25cm}}
		%%		 Optimal schedule length is 23.
	\end{figure}

	\begin{table}[!tb]
	\begin{center}
		\begin{tabular}{|c|c|c|c|}
			\hline
			Task & Memory Req & Comm Time & Comp Time\\ \hline 
%%			\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
%%			&=Comm Volume && \\ \hline
			A & 3 & 3 & 2\\ \hline
			B & 1 & 1 &  6\\ \hline
			C & 4 & 4 & 6\\ \hline
			D & 5 & 5 & 1\\ \hline
		\end{tabular}
		\caption{\label{tab:dynamicSelectionExample} A task set for dynamic schedules.\vspace*{-0.75cm}}
	\end{center}
	\end{table}

@@

	\subsection{Static Order with Dynamic Correction}

	In this class of strategy, we precompute the order of tasks based on some criterion and then follow this ordering as much as possible. But when the communication resource is idle because the memory requirement of the next task is too high, then we select a task  with a dynamic strategy. After a task is dynamically selected, we update the remaining order without this task. This class of strategy takes advantage of static information in the form of precomputed order and dynamic correction to minimize the idle time due to memory constraint.

@@

	%%\begin{figure}[htb]
	%%\includegraphics[scale=0.05]{Figs/staticOrderDynamiccorrectionSchedules}
	%%\caption{Task set and schedules where orders are precomputed based on different heuristics.}
	%%\label{fig:staticOrderDynamiccorrectionExample}
	%%\end{figure}
\begin{figure}[!htb]
	\newcommand{\taskA}[2][0]{\task[#1]{$A$}{#2}{4}{1}{cyan}}
	\newcommand{\taskB}[2][0]{\task[#1]{$B$}{#2}{2}{6}{cyan!50!black}}
	\newcommand{\taskC}[2][0]{\task[#1]{$C$}{#2}{8}{8}{blue!40!white}}
	\newcommand{\taskD}[2][0]{\task[#1]{$D$}{#2}{5}{4}{blue!70!white}}
	\newcommand{\taskE}[2][0]{\task[#1]{$E$}{#2}{3}{2}{blue}}
	\vspace*{-0.35cm}
	\begin{tikzpicture}[yscale=0.5,xscale=0.2]
	\begin{scope}
	\schedule{35}{OOLCMR}{2,8,12,15,17,25,33}
	\taskB{0}
	\taskD[1]{2}
	\taskA{8}
	\taskE{12}
	\taskC{17}
	\end{scope}
	\begin{scope}[yshift=-2.75cm]
	%%		\schedule{36}{OOSCMR}{2,5,8,11,14,18,26,34}
	%%		\taskB{0}
	%%		\taskE[3]{2}
	%%		\taskA[1]{5}
	%%		\taskD{9}
	%%		\taskC{18}
	%%		\end{scope}
	\schedule{36}{OOSCMR}{2,5,8,11,15,19,27,35}
	\taskB{0}
	\taskE[3]{2}
	\taskA[1]{5}
	\taskD{10}
	\taskC{19}
	\end{scope}
	\begin{scope}[yshift=-5.5cm]
	\schedule{36}{OOMAMR}{2,8,11,14,17,25,33}
	\taskB{0}
	\taskD[1]{2}
	\taskE[1]{8}
	\taskA{12}
	\taskC{17}
	\end{scope}
	%%		\begin{scope}[yshift=+2.75cm]
	%%		\schedule{40}{OMIM}{2,8,16,24,29,33,38}
	%%		\taskB{0}
	%%		\taskC{8}
	%%		\taskD{24}
	%%		\taskA{29}
	%%		\taskE{33}
	%%		\end{scope}

	\end{tikzpicture}
	\caption{ \label{fig:staticOrderDynamiccorrectionExample} Different static order dynamic correction heuristic schedules for a task set of Table~\ref{tab:staticOrderDynamiccorrectionExample} with a memory capacity of 9. The $OMIM$ order is $BCDEA$.\vspace*{-0.5cm}}
\end{figure}

\begin{table}[!htb]
	\begin{center}
		\begin{tabular}{|c|c|c|c|}
			\hline
			Task & Memory Req & Comm Time & Comp Time\\ \hline 
%%			\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
%%			&=Comm Volume && \\ \hline
			A & 4 & 4 &  1\\ \hline
			B & 2 & 2 & 6\\ \hline
			C & 8 & 8 & 8\\ \hline
			D & 5 & 5 & 4\\ \hline
			E & 3 & 3 & 2\\ \hline
		\end{tabular}
		\caption{\label{tab:staticOrderDynamiccorrectionExample} A task set for static order dynamic correction schedules.\vspace*{-0.75cm}}
	\end{center}
\end{table}

	As previously, we propose on Table~\ref{tab:staticOrderDynamiccorrectionExample} an example instance for this class of strategies, and provide on Figure~\ref{fig:staticOrderDynamiccorrectionExample} the corresponding schedules.

@@

			Based on the definition of proposed strategies and the optimality of Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}, we discuss the scenarios which might be more favorable for each heuristic in Table~\ref{tab:heuristicsAndFavorableScenarios}. This allows programmers to use appropriate strategies to maximize communication-computation overlap for their applications. In this table, ``moderate memory capacity'' refers to the case where memory is constrained, but close to the maximal memory requirement of the $OMIM$ schedule.

		\begin{table}[!h]
			\begin{center}
			\begin{tabular}{|c|p{6.5cm}|}
				\hline
				\textbf{Heuristic} & \textbf{\hspace{2cm}Favorable Situation} \\ \hline
				$OOSIM$ & Memory capacity is not a restriction (\textcolor{green}{Optimal}) \\ \hline
				$IOCMS$ & Memory capacity is not a restriction and tasks are compute intensive (\textcolor{green}{Optimal}) \\ \hline
				$DOCPS$ & Memory capacity is not a restriction and tasks are communication intensive (\textcolor{green}{Optimal}) \\ \hline
				$IOCCS$ & Moderate memory capacity and most tasks are highly compute intensive \\ \hline
				$DOCCS$ & Moderate memory capacity and most tasks are highly communication intensive \\ \hline
				$LCMR$ & Limited memory capacity and significant percentage of tasks with large communication times are compute intensive\\ \hline
				$SCMR$ & Limited memory capacity and significant percentage of tasks with small communication times are compute intensive\\ \hline
				$MAMR$ & Limited memory capacity and significant percentage of all types of tasks\\ \hline
				$OOLCMR$ & Moderate memory capacity and significant percentage of slightly communication intensive tasks have large communication times\\ \hline
				$OOSCMR$ & Moderate memory capacity and significant percentage of slightly communication intensive tasks have small communication times\\ \hline
				$OOMAMR$ & Moderate memory capacity and significant percentage of all types of tasks \\ \hline
			\end{tabular}\caption{~\label{tab:heuristicsAndFavorableScenarios}Heuristics and their favorable scenarios.\vspace*{-0.75cm}}
			\end{center}
		\end{table}

@@

		%%		\todo[inline]{More about tiling}
		Both applications mainly perform two types of computations, tensor transpose and tensor contraction. HF expects to specify a tile size and we set it to 100, so that each core can be used efficiently. CCSD automatically determines tile sizes at different program points based on the input molecules. Hence, HF operates on almost homogeneous tiles while CCSD uses more heterogeneous tiles.

		\subsection{Workload Characteristics}
		\begin{figure}[htb]
			\vspace*{-0.3cm}
			\begin{center}
				\includegraphics[scale=0.48]{./results/application_properties.pdf}
				\vspace*{-0.9cm}\caption{HF and CCSD tasks characteristics.\vspace*{-0.35cm}}
				\label{fig:ApplicationProperties}
			\end{center}
		\end{figure}
			\begin{figure*}[!t]
			\includegraphics[scale=0.5]{./results/ratio_to_optimal_selected_hf.pdf}
			\vspace*{-0.85cm}\caption{Comparison of different heuristics for HF with $m_c=176KB$.			\vspace*{-0.35cm}}
			\label{fig:ratio_to_optimal_hf}
		\end{figure*}

		\noindent To get more insights into the considered workloads, we provide  information about the instances we consider in Figure~\ref{fig:ApplicationProperties}. For each instance, we compute the sum of communication times (\textit{sum comm}) and the sum of computation times  (\textit{sum comp}), and normalize it relatively to the $OMIM$ value. Figure~\ref{fig:ApplicationProperties} also shows the maximum of both values, which is a lower bound on the possible makespan of a schedule for this instance, and their sum, which is an upper bound: this represents the makespan of the sequential schedule, obtained with zero overlap between computation and communication. We can see that HF is a communication intensive application and at most 20\% overlap can be expected in the best scenario. On the other hand, in the CCSD workload, communications and computations are almost evenly distributed and a more significant overlap is possible.

@@

		The CCSD application operates on tasks of different sizes, hence different heuristics exhibit distinct behaviors even at minimum memory capacity $m_c$. Heterogeneity favors dynamic strategies, as can be seen by the fact that both dynamic and static order with dynamic correction based strategies perform better than static based strategies. Similar to HF, static order with dynamic correction based strategies outperform others as memory capacity becomes moderate.

\begin{figure}[htb]
	\includegraphics[scale=0.5]{./results/ratio_to_optimal_ccsd-best.pdf}
	\vspace*{-0.25cm}\caption{Best variants of all categories for CCSD.\vspace*{-0.25cm}}
	\label{fig:ratio_to_optimal_best_ccsd}
\end{figure}
